Planetary perturbers: Flaring star-planet interactions in Kepler and TESS

### 2023-10-03 src/tex/ms.tex
[showyourwork] overleaf sync

The methods for light curve de-trending and flare finding in Kepler and TESS light curves, as well as the Anderson-Darling test, are the same as detailed in~\citet{ilin2022searching}. We briefly recap the techniques in Sections~\ref{sec:methods:flaresearch} and \ref{sec:methods:adtest}. The expected power of SPI depends on the relative velocity between the planet and the magnetic field strength in its orbit, the derivation of which we explain in 
Sections~\ref{sec:methods:relvel} and \ref{sec:methods:bfield}, respectively. We can then combine them with stellar radius, planetary radius, and semi-major axis to estimate the power of SPI. We use the scaling laws for two different magnetic SPI models, which we introduce in Section~\ref{sec:methods:pspi}.

@@

\subsection{Custom Anderson-Darling test}
\label{sec:methods:adtest}
For each star-planet system with three or more flares in its light curves, we test for deviations from a random distribution of flares with orbital phase. We use the same customized Anderson-Darling test as in \cite{ilin2022searching}, which is more sensitive to deviations at both ends of the distribution than the widely used Kolmogorov-Smirnov test~\citep{feigelson2012modern}. In brief, we first take the number of flares observed in the TESS and Kepler data to calculate a base flare rate for each light curve. We then calculate how often each orbital phase has been observed, also for each light curve separately, to ensure that the different detection thresholds for TESS and Kepler do not bias the base rate. With the phase coverage and base flare rate combined, we can tell how many flares we would expect to see in any given phase bin if flares were randomly distributed. We then aggregate this number of expected flares per time bin into one expected distribution by summing over all available light curves. As a last step, we compare the expected distribution to the observed one using an Anderson-Darling test. \textcolor{red}{For this, we sample 10000 times from the expected distribution the same number of flares we observe in each system, e.g., for HIP 67522, we sample 12 flares distributed according to the blue curve in Fig.~\ref{fig:cumdist_transiting}. For each of the 10000 samples, we compute the Anderson-Darling statistic $A^2$, which yields a distribution of $A^2$ values. In a last step, we compute $A^2$ for the observed distribution, and compute the $p$-value.} We repeat this test with different phase offsets to account for remaining biases in sensitivity of the test at different phases.

@@

\subsection{Stellar magnetic fields}
\label{sec:methods:bfield}
We derive the average magnetic field strength $B$ from $R$o using the empirical relation derived in \citet{reiners2022magnetism},  Table 2, in the unsaturated and saturated regimes, respectively:

@@

First, we estimate the power of SPI generated by the stretch-and-break mechanism using the scaling relations in~\cite{lanza2012starplanet} (their Eqn. 45), where $P_{\rm spi,sb}$ is proportional to

\begin{equation}
    P_{\rm spi,sb} \sim B_{\rm p}^{2/3} R_{\rm p}^2 B_*^{4/3} R_*^4 v_{\rm rel} a^{-4} .
\end{equation}

$B_{\rm p}$ and $B_*$ are the planetary and stellar field strengths at their poles, respectively; $v_{\rm rel}$ is the relative velocity between the stellar rotation and the planet's orbit at the semi-major axis $a$; and $R_*$ and $R_{\rm p}$ are the stellar and planetary radii, respectively. Since the planetary fields are unknown, we ignore them in the calculation. We also do not know the stellar field strength at the pole for the vast majority of the stars, so we adopt the surface field strength~(Section~\ref{sec:methods:bfield}) as a proxy.

@@

\begin{equation}
    P_{\rm spi,sb0} \sim R_{\rm p}^2 R_*^4 B_*^2 v_{\rm rel}    a^{-4} 
\end{equation}

@@

\begin{equation}
    P_{\rm spi,aw} \sim B_{\rm p}^{2/3} R_{p}^2  R_*^2  B_*^{1/3}  v_{\rm rel}^2 a^{-2} \rho_*^{1/2} \sin^2 \theta
\end{equation}

@@

And fourth, we also consider the case of an unmagnetized planet\textcolor{red}{ where the magnetospheric radius is set equal to the planetary radius}, in which case Eq.~8 in \cite{kavanagh2022radio} reduces to

\begin{equation}
    P_{\rm spi,aw0} \sim R_{\rm p}^2 R_*^4 B_*  v_{\rm rel}^2 a^{-4} \rho_*^{1/2} \sin^2 \theta
\end{equation}

@@

\begin{table*}
\footnotesize
\movetableright=-20mm
\script{paper_main_table.py}
\caption{Flaring star-planet interaction. $Ro$, $B$, and $v_{\rm rel}$, are derived from literature values (Table~\ref{tab:maintable_lit}). $P_{\rm spi,xx}$ stands for the power of stretch-and-break ($sb$) and Alfv\'en wing ($aw$) interaction mechanisms, assuming the planet has a magnetic field strength of $1\,$G. $P_{\rm spi,xx0}$ is the same, but assuming an unmagnetized planet. All powers are normalized to AU Mic. The $p$-value of the Anderson-Darling test is lower when the system shows a stronger clustering of flares in orbital phase.}
\input{output/table_der_vals.tex}
    \label{tab:maintable_der}

@@

This relation links the mean AS size with the maximum large-scale radial magnetic field ($B_{\rm r}^{\rm max}$) obtained in Zeeman-Doppler Imaging reconstructions (\citealt{donati2006surprising, morin2008largescale, fares2009magnetic, alvarado-gomez2015activity, hussain2016spectropolarimetric, kochukhov2020mapping}). The relation was derived from 3D~MHD stellar wind simulations of cool main sequence stars (F, G, K, and M), which used observed ZDI maps as boundary condition for the stellar radial magnetic field (for more details on the models see \citealt{chebly2023numerical}). However, the stellar magnetic field~($B$, Table~\ref{tab:maintable_der}), derived from the Rossby number (Eq.~\ref{Eq3}, \ref{Eq4}), corresponds to the total, unsigned magnetic field strength on the stellar surface, so that it is necessary to estimate how $B_{\rm r}^{\rm max}$ is related to $B$.

>>>>

The methods for light curve de-trending and flare finding in Kepler and TESS light curves, as well as the Anderson-Darling test, are the same as detailed in~\citet{ilin2022searching}. We briefly recap the techniques in Sections~\ref{sec:methods:flaresearch} and \ref{sec:methods:adtest}. The expected power of SPI depends on the relative velocity between the planet and the magnetic field strength in its orbit, the derivation of which we explain in 
Sections~\ref{sec:methods:relvel} and \ref{sec:methods:bfield}, respectively. We can then combine them with \textcolor{red}{the} stellar radius, planetary radius, and semi-major axis to estimate the power of SPI. We use the scaling laws for two different magnetic SPI models, which we introduce in Section~\ref{sec:methods:pspi}.

@@

\subsection{Custom Anderson-Darling test}
\label{sec:methods:adtest}
For each star-planet system with three or more flares in its light curves, we test for deviations from a random distribution of flares with orbital phase. We use the same customized Anderson-Darling test as in \cite{ilin2022searching}, which is more sensitive to deviations at both ends of the distribution than the widely used Kolmogorov-Smirnov test~\citep{feigelson2012modern}. In \textcolor{red}{short}, we first take the number of flares observed in the TESS and Kepler data to calculate a base flare rate for each light curve. We then calculate how often each orbital phase has been observed, also for each light curve separately, to ensure that the different detection thresholds for TESS and Kepler do not bias the base rate. With the phase coverage and base flare rate combined, we can tell how many flares we would expect to see in any given phase bin if flares were randomly distributed. We then aggregate this number of expected flares per time bin into one expected distribution by summing over all available light curves. As a last step, we compare the expected distribution to the observed one using an Anderson-Darling test. \textcolor{red}{For this, we sample 10000 times from the expected distribution the same number of flares we observe in each system, e.g., for HIP 67522, we sample 12 flares distributed according to the blue curve in Fig.~\ref{fig:cumdist_transiting}. For each of the 10000 samples, we compute the Anderson-Darling statistic $A^2$, which yields a distribution of $A^2$ values. In a last step, we compute $A^2$ for the observed distribution, and compute the $p$-value.} We repeat this test with different phase offsets to account for remaining biases in sensitivity of the test at different phases.

@@

\subsection{Stellar magnetic fields}
\label{sec:methods:bfield}
We derive the average magnetic field strength $B$ from \textcolor{red}{Rossby number} $R$o using the empirical relation derived in \citet{reiners2022magnetism},  Table 2, in the unsaturated and saturated regimes, respectively:

@@

First, we estimate the power of SPI generated by the stretch-and-break mechanism using the scaling relations in~\cite{lanza2012starplanet}. \textcolor{red}{We adopt their Eqn. 45 for non-linear and axisymmetric magnetic fields, because the non-linear formulation does not assume a constant force-free parameter, and because only axisymmetric fields are able to produce modulation of flaring activity in phase with the planetary orbit. We note that the axisymmetry of stellar magnetic fields varies across stars~\citep[e.g.,][]{donati2009magnetic, donati2008largescale, morin2008largescale, morin2010largescale}, and also with the stellar activity cycle of individual stars~\citep[e.g.,][]{borosaikia2016solarlike, lehmann2021identifying} In Eqn. 45,} $P_{\rm spi,sb}$ is proportional to

\begin{equation}
    P_{\rm spi,sb} \propto B_{\rm p}^{2/3} R_{\rm p}^2 B_*^{4/3} R_*^{(n+11) / 3} v_{\rm rel} a^{-(n+11) / 3},
\end{equation}

\textcolor{red}{ignoring factors of the order of 1.}  $B_{\rm p}$ and $B_*$ are the planetary and stellar field strengths at their poles, respectively; $v_{\rm rel}$ is the relative velocity between the stellar rotation and the planet's orbit at the semi-major axis $a$; and $R_*$ and $R_{\rm p}$ are the stellar and planetary radii, respectively. \textcolor{red}{We choose $n=1$ to quantify the strongest possible dependence of the dissipated power on stellar radius and star-planet distance.} Since the planetary fields are unknown, we ignore them in the calculation. We also do not know the stellar field strength at the pole for the vast majority of the stars, so we adopt \textcolor{red}{an average value of 15 per cent of the} surface field strength~(see Section~\ref{sec:methods:bfield}) as a proxy.

@@

\begin{equation}
    P_{\rm spi,sb0} \propto R_{\rm p}^2 R_*^{(n+11) / 3} B_*^2 v_{\rm rel}    a^{-(n+11) / 3} 
\end{equation}

@@

\begin{equation}
    P_{\rm spi,aw} \propto B_{\rm p}^{2/3} R_{p}^2  R_*^2  B_*^{1/3}  v_{\rm rel}^2 a^{-2} \rho_*^{1/2} \sin^2 \theta
\end{equation}

@@

And fourth, we also consider the case of an unmagnetized planet, \textcolor{red}{ where the magnetospheric radius is set equal to the planetary radius}, in which case Eq.~8 in \cite{kavanagh2022radio} reduces to

\begin{equation}
    P_{\rm spi,aw0} \propto R_{\rm p}^2 R_*^4 B_*  v_{\rm rel}^2 a^{-4} \rho_*^{1/2} \sin^2 \theta
\end{equation}

@@

\begin{table*}
\footnotesize
\movetableright=-20mm
\script{paper_main_table.py}
\caption{Flaring star-planet interaction. $R$o, $B$, and $v_{\rm rel}$, are derived from literature values (Table~\ref{tab:maintable_lit}). $P_{\rm spi,xx}$ stands for the power of stretch-and-break ($sb$) and Alfv\'en wing ($aw$) interaction mechanisms, assuming the planet has a magnetic field strength of $1\,$G. $P_{\rm spi,xx0}$ is the same, but assuming an unmagnetized planet. All powers are normalized to AU Mic. The $p$-value of the Anderson-Darling test is lower when the system shows a stronger clustering of flares in orbital phase.}
\input{output/table_der_vals.tex}
    \label{tab:maintable_der}

@@

This relation links the mean AS size with the maximum large-scale radial magnetic field ($B_{\rm r}^{\rm max}$) obtained in Zeeman-Doppler Imaging reconstructions (\citealt{donati2006surprising, morin2008largescale, fares2009magnetic, alvarado-gomez2015activity, hussain2016spectropolarimetric, kochukhov2020mapping}). The relation was derived from 3D~MHD stellar wind simulations of cool main sequence stars (F, G, K, and M), which used observed ZDI maps as boundary condition for the stellar radial magnetic field (for more details on the models see \citealt{chebly2023numerical}). However, the stellar magnetic field~($B$, Table~\ref{tab:maintable_der}), derived from the Rossby number (Eq.~\ref{Eq3}, \ref{Eq4}), corresponds to the total, unsigned magnetic field strength on the stellar surface so that it is necessary to estimate how $B_{\rm r}^{\rm max}$ is related to $B$.

\textcolor{red}{It is important to note that in \citet{chebly2023numerical} they assume a consistent Poynting flux per unit magnetic field strength for all stars, using the value of the Sun as a reference. The size of the Alfvén surface depends strongly on this assumption. However, any variation in the value between different stars would lead to deviations from the law proposed in \citet{chebly2023numerical}, resulting in corresponding changes in the dimensions of the Alfvén surface.
}
